---
title: "Whole Genome Coverage & Genome Filtering for S. epidermidis Serial Samples"
author: "Nicholas F. Brazeau"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    highlight: textmate
    theme: lumen
    toc: yes
    toc_float: yes
    toc_depth: 2
    fig_caption: yes
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```
```{r imports}
library(tidyverse)
library(RColorBrewer)
library(DT)

remotes::install_github("nickbrazeau/NFBtools", ref = "6b441643c74336b636282fcff2b1b51ca3eb0094")
library(NFBtools) # just using the basic functionality of bamCov.R 

```
# Background
Over their course of treatment at UNC, the patient developed multi-drug resistant _S. epidermidis_. I am currently blinded to timing and resistance phenotype data but am attempting to identify putative drug resistance mutations. 

## Goal 
**Overall**: To detect putative drug resistance mutation(s) in clinical isolates of _S. epidermidis_ from a patient with phenotypic resistance.   
**Immediate**: Quality control of samples and determine read depth.




## FlagStats
```{r flagstat, results='asis'}
flagsumfiles <- dir(path="~/Documents/MountPoints/mountedScratchLL/Projects/Sepi_Res_CaseStudy/SumSTATsandQC/FlagStats", full.names = TRUE)

readflag <- function(path){
  out <- read_tsv(path, col_names = F)
  sample <- stringr::str_split_fixed(string=basename(path), pattern="_", n=2)[1]
  out <- as.data.frame(cbind(sample, out))
  colnames(out) <- c("sample", "summary_stat")
  return(out)
}

flagsumfiles <- lapply(flagsumfiles, readflag)
flagsumfiles <- do.call("rbind", flagsumfiles)

DT::datatable(flagsumfiles, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('csv')))


```


## Callable Loci Summary
Output is based on a mimimum base quality of 20, minimum mapping quality of 10, and a minimum depht of 4 (these are all > criterion).  
  
```{r calloci, results='asis'}
calllocisumfiles <- dir(path="~/Documents/MountPoints/mountedScratchLL/Projects/Sepi_Res_CaseStudy/SumSTATsandQC/CallableLoci/", full.names = TRUE)
calllocisumfiles <- calllocisumfiles[grepl(pattern = "_summarytable.txt", x = basename(calllocisumfiles))]
readcallloci <- function(path){
  out <- read_tsv(path, col_names = T)
  sample <- stringr::str_split_fixed(string=basename(path), pattern="_", n=2)[1]
  out <- as.data.frame(cbind(sample, out))
  return(out)
}

calllocisumfiles <- lapply(calllocisumfiles, readcallloci)
calllocisumfiles <- do.call("rbind", calllocisumfiles)

DT::datatable(calllocisumfiles, extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('csv')))


```

## Genomic Coverage by Sample
Presented is the overall Genomic Coverage and a Summary Table by Sample. 
```{r genomcov, fig.width=12, fig.height=12, dpi=500}

#------------------------
# read in
#------------------------
bamcovs <- NFBtools::bedtoolsgenomecov2bamCov(gencovdir = "~/Documents/MountPoints/mountedScratchLL/Projects/Sepi_Res_CaseStudy/SumSTATsandQC/coverage/data/")

#------------------------
# genomic summary plot
#------------------------
genomcovplot <- NFBtools::bamCov2OverallPercCov(bamcovs)


#------------------------
# genomic summary table
#------------------------
genomsummarydepth <- do.call("rbind", lapply(bamcovs, function(x){x[["genomsummarydepth"]]})) %>% 
  dplyr::mutate_if(is.numeric, round, 2)

#https://stackoverflow.com/questions/11774703/adding-text-to-a-grid-table-plot/11775211#11775211
genomtable <- gridExtra::tableGrob(genomsummarydepth, rows = NULL)
title <- grid::textGrob("Summary of Coverage Depth by Sample", gp=grid::gpar(fontfamily="Arial", fontsize=14, fontface="bold"))
padding <- unit(5,"mm")
genomtable <- gtable::gtable_add_rows(
        genomtable,
        heights = grid::grobHeight(title) + padding,
        pos = 0)
genomtable <- gtable::gtable_add_grob(
        genomtable,
        list(title),
        1, 1, 1, ncol(genomtable))

grobs = list(genomcovplot, genomtable)
gridExtra::grid.arrange(grobs = grobs, layout_matrix = rbind(c(1),
                                                             c(1),
                                                             c(2)
                                                             )
                        )


jpeg(paste0("~/Documents/GitHub/Sepi_Res_CaseStudy/Ranalysis/figures/SFig1-Genomic_Coverage.jpg"), units = "in", width = 8, height = 11, res = 200)
gridExtra::grid.arrange(grobs = grobs, layout_matrix = rbind(c(1),
                                                             c(1),
                                                             c(2)
                                                             )
                        )
graphics.off()


```

## Sample-Level Genomic Coverage
```{r smplcov, fig.width=18, fig.height=12, results='hide', dpi=500}


NFBtools::bamCov2SmplRaster(input = bamcovs)



```

## Windowed Coverage
```{r windowcovbysmpl}

do.call("rbind.data.frame", purrr::map(purrr::map(bamcovs, "windowcov"), "windowcovdf")) %>% 
  tibble::as_tibble(.) %>% 
  DT::datatable(., extensions = 'Buttons', options = list(dom = 'Bfrtip',buttons = c('csv'), pageLength = 100))

```

```{r}

saveRDS(bamcovs, file = "~/Documents/GitHub/Sepi_Res_CaseStudy/Ranalysis/data/Sepi_coverage.RDS")

```
