---
title: "Whole Genome INDEL Analysis for Sepi Resistance"
author: "Nicholas F. Brazeau"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  rmarkdown::html_document:
    highlight: textmate
    theme: lumen
    toc: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, fig.width = 10, fig.height = 7)
```

```{r}
#............
# imports
#..........
library(vcfR)
library(tidyverse)
library(RColorBrewer)
library(grid)
library(DT)
devtools::install_github("IDEELResearch/vcfRmanip")
library(vcfRmanip)


```

# Goal 
**Overall**: To detect putative drug resistance mutation(s) in clinical isolates of _S. epidermidis_ from a patient with phenotypic resistance.   
**Immediate**: QC and Analysis of Segregating Sites and putative Drug-rx mutations of structural variants.

## Background

Over their course of treatment, a patient developed multi-drug resistant _S. epidermidis_. We are attempting to identify putative drug resistance mutations in line with the phenotypic data. **The purpose of this script is to analyze the potential Structural Variants among the clinical isolates**. 

 Thus far, I have done the following: 

1. Aligned with `bwa-mem` with default parameters to fasta from [NCBI](https://www.ncbi.nlm.nih.gov/genome/155?genome_assembly_id=299299)
2. Deduplicated and mate-fixed PE reads (python modules)
3. Variant called with `smoove/lumpy`
4. Used `duphold` to subset to more likely duplications and deletions
5. Kara Moser used `mummer` to make inversion and BND bed

```{r}
# VCF IMPORT
lumpyvcf <- vcfR::read.vcfR("~/Documents/GitHub/Sepi_Res_CaseStudy/smoothlumpy/bams/sepi_clincases-smoove.genotyped.vcf.gz", verbose = F)

```

Overall, there appears to be `r nrow(lumpyvcf@gt)` structural variants identifed by smoove. Going to investigate these for association with the clinical phenotype.


## Variant Filtration
### Seg Sites
```{r}

lumpyvcf.seg <- vcfRmanip::vcfR2segsites_gt(lumpyvcf)
lumpyvcf.seg
```


### Mask Low Complexity Regions
```{r}
repeatregions <- readRDS(file = "~/Documents/GitHub/Sepi_Res_CaseStudy/analyze_sepi_cc/data/derived_data/lowcomplexity_repeat_regions.RDS")

# mask lumpy

bpbuffer <- 50

repeatregions <- repeatregions %>% 
  dplyr::rename(start = start_orient,
                end = end_orient) %>% 
  dplyr::mutate(
    seqname = "Chromosome",
    geneid = 1:nrow(.),
    start = start - bpbuffer,
    end = end + bpbuffer 
  ) 


lumpyvcf.seg.masked <- vcfRmanip::vcffilter_ChromPos(vcfRobject = lumpyvcf.seg,
                                                 chromposbed = repeatregions)



```


### Apply Filter for Dels and Dups
Following the suggestions from `duphold` [manuscript](https://academic.oup.com/gigascience/article/8/4/giz040/5477467) that for deletions
a DHFFC < 0.7 and for duplicates a DHBFC > 1.3 lead to accurate calls.

```{r}

lumpyvcf.seg.masked.long <- lumpyvcf.seg.masked %>% 
  vcfR::vcfR2tidy(.)

lumpyvcf.seg.masked.long <- dplyr::left_join(lumpyvcf.seg.masked.long$fix, lumpyvcf.seg.masked.long$gt, by = c("ChromKey", "POS"))
```

#### Duplications

```{r}

passDUPs.lumpyvcf.seg.masked.long <- lumpyvcf.seg.masked.long %>% 
  dplyr::filter(ALT == "<DUP>" & gt_DHBFC > 1.3)


passDUPs.gtmat <- passDUPs.lumpyvcf.seg.masked.long %>% 
  dplyr::select(c("CHROM", "POS", "ID", "SVTYPE", "SVLEN", "END", "Indiv", "gt_GT")) %>% 
  tidyr::spread(., key = "Indiv", value = "gt_GT")

```

```{r, results='asis'}
DT::datatable(passDUPs.gtmat, extensions='Buttons',
               options = list(
              searching = T,
              pageLength = 5,
              dom = 'Bfrtip', 
              buttons = c('csv')))

```


#### Deletions
```{r}


passDELs.lumpyvcf.seg.masked.long <- lumpyvcf.seg.masked.long %>% 
  dplyr::filter(ALT == "<DEL>" & gt_DHFFC < 0.7)


passDELs.gtmat <- passDELs.lumpyvcf.seg.masked.long %>% 
  dplyr::select(c("CHROM", "POS", "ID", "SVTYPE", "SVLEN", "END", "Indiv", "gt_GT")) %>%
  tidyr::spread(., key = "Indiv", value = "gt_GT")

```

```{r, results='asis'}
DT::datatable(passDELs.gtmat, extensions='Buttons',
               options = list(
              searching = T,
              pageLength = 5,
              dom = 'Bfrtip', 
              buttons = c('csv')))

```


### Apply Filter for BND and Inv
Compare complex rearrangement and inversion calls between lumpy and mummer. Only take those that match. Note, no segregating inversion found. 

```{r}

mummerbed <- readr::read_tsv("~/Documents/GitHub/Sepi_Res_CaseStudy/analyze_sepi_cc/data/derived_data/all.potential-inversions.bed", col_names = F) %>% 
  magrittr::set_colnames(c("CHROM", "POS", "END", "SVLEN", "Indiv")) %>% 
  dplyr::mutate(Indiv = stringr::str_split_fixed(Indiv, "_", n=2)[,1])

mummerbed.chrompos <- mummerbed  %>% 
  dplyr::select(c("CHROM", "POS", "END", "SVLEN")) %>% 
  dplyr::filter(!duplicated(.)) %>% 
  dplyr::arrange(POS) %>% 
  dplyr::mutate(y = 1:nrow(.))



BND <- lumpyvcf.seg.masked.long %>% 
  dplyr::select(c("CHROM", "POS", "ID", "SVTYPE", "SVLEN", "END", "Indiv", "gt_GT")) %>%
  dplyr::mutate(Indiv = tolower(Indiv)) %>% 
  dplyr::filter(SVTYPE == "BND")

BND.chrompos <- BND %>% 
  dplyr::select(c("CHROM", "POS", "END", "SVLEN")) %>% 
  dplyr::filter(!duplicated(.)) %>% 
  dplyr::mutate(y = 1:nrow(.))



BND.chrompos.adp <- data.frame(start = 720794, end = 1518329 )

```

##### Visualize Mummer "Dot Plot" For All Potential Inversions
```{r, results='asis'}
ggplot() + 
  geom_rect(data = mummerbed.chrompos, aes(xmin = POS, xmax = END, ymin = y+0.25, ymax=y+0.75), color = "black") +
  geom_rect(data = BND.chrompos.adp, aes(xmin = start, xmax = end, ymin=0, ymax=Inf), fill = "red", alpha = 0.2)
  
  


```

**Based on visualization, we do not find support this complex breakend.**

## Analyzing All Indels
### GT Mat for Raw Segs
```{r, results='asis'}

rawchrompos <- data.frame(CHROM = vcfR::getCHROM(lumpyvcf.seg), 
                          POS = vcfR::getPOS(lumpyvcf.seg), 
                          SVTYPE = vcfR::extract.info(lumpyvcf.seg, "SVTYPE"),
                          SVLEN = vcfR::extract.info(lumpyvcf.seg, "SVLEN"),
                          END = vcfR::extract.info(lumpyvcf.seg, "END")
                          )

rawgt <- tibble::as_tibble( vcfR::extract.gt(lumpyvcf.seg, element = "GT") ) %>%
  dplyr::select(c("Sepi01", "Sepi02", "Sepi03", "Sepi04", "Sepi05a", "Sepi05b", "Sepi06", "Sepi07")) # reorder
  

DT::datatable(cbind.data.frame(rawchrompos, rawgt), 
              extensions='Buttons',
               options = list(
              searching = T,
              pageLength = 5,
              dom = 'Bfrtip', 
              buttons = c('csv')))

```


### GT Mat for Filtered Segegrating SVs


```{r, results='asis'}
passSV <- rbind.data.frame(passDELs.gtmat, passDUPs.gtmat) 
  DT::datatable(passSV, 
                extensions='Buttons',
                options = list(
                  searching = T,
                  pageLength = 5,
                  dom = 'Bfrtip', 
                  buttons = c('csv')))

```


```{r}
save(lumpyvcf.seg,
     lumpyvcf.seg.masked,
     passSV,
     file = "~/Documents/GitHub/Sepi_Res_CaseStudy/analyze_sepi_cc/data/lumpy_raw_filt.RDA")


```

